### 2010-03-24 Paper/Paper.tex
minor

% -*- TeX-PDF-mode: t -*-

% -----------------------------------------------
% Template for ISMIR 2010
% (based on earlier ISMIR templates)
% -----------------------------------------------

\documentclass{article}
\usepackage{ismir2010,amsmath,cite}
\usepackage{graphicx}
\usepackage{url}
\usepackage{algorithm,algorithmic}

% added by tbm
\usepackage{subfloat,subfig}
\setlength{\abovecaptionskip}{-3pt}
\setlength{\belowcaptionskip}{6pt}

\usepackage{booktabs}

\DeclareMathOperator*{\dist}{dist}
\DeclareMathOperator*{\size}{size}

% Title.
% ------
\title{Clustering Beat-Chroma Patterns\\ in a Large Music Database}

% Single address
% To use with only one author or several with the same address
% ---------------
%\oneauthor
% {Names should be omitted for double-blind reviewing}
% {Affiliations should be omitted for double-blind reviewing}

% Two addresses
% --------------
%\twoauthors
%  {First author} {School \\ Department}
%  {Second author} {Company \\ Address}

% Three addresses
% --------------
\threeauthors
  {First author} {Affiliation1 \\ {\tt author1@ismir.edu}}
  {Second author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
  {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
% what order do we use? Thierry, Ron, Dan?  Dan, Thierry, Ron?

% Q: Why are there long spaces after "e.g." or "i.e."?
% A: LaTeX interprets a dot and a blank as the end of a sentence. To
%    prevent this, use "e.g.\ " and "i.e.\ " or "e.g.~" and "i.e.~".
\newcommand{\ie}{i.e.~}
\newcommand{\Ie}{I.e.~}
\newcommand{\eg}{e.g.~}
\newcommand{\Eg}{E.g.~}

\begin{document}
%
\maketitle
%
\begin{abstract}
A musical style or genre implies a set of common conventions
and patterns combined and deployed in different ways to make
individual musical pieces; for instance, most would agree that
contemporary pop music is assembled from a relatively small
palette of harmonic and melodic patterns.  The purpose of this
paper is to use a database of tens of thousands of songs
in combination with a compact representation of melodic-harmonic
content, the beat-synchronous chromagram, and data-mining
tools (clustering) to attempt to explicitly catalog this palette --
at least within the limitations of the beat-chroma representation.
We use online $k$-means clustering to summarize 3.7 million
4-beat bars in a codebook of a few hundred prototypes.
By measuring how accurately such a quantized codebook
can reconstruct the original data, we can quantify the degree
of diversity (distortion as a function of codebook size) and
temporal structure (\ie the advantage gained
by joint quantizing multiple frames) in this music.  The most
popular codewords themselves reveal the common chords
used in the music.  Finally, the quantized representation of
music can be used for music retrieval tasks such as artist
and genre classification, and identifying songs that are
similar in terms of their melodic-harmonic content.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}
The availability of very large collections of music audio present
many interesting research opportunities.  Given millions of examples
from a single, broad class (\eg contemporary commercial pop music),
can we infer anything about the underlying structure and common
features of this class?  This paper describes our work in this direction.

What are the common features of pop music?  There are conventions
of subject matter, instrumentation, form, rhythm, harmony, and melody, among
others.  Our interest here is in the tonal content of the music -- \ie the
harmony and melody.  As a computationally convenient proxy for a richer
description of the tonal content of audio, we use the popular Chroma
representation, which collapses an acoustic spectrum into a 12-dimensional
description, with one bin for each semitone of the western musical octave.
In addition, we simplify the time axis of our representation to take advantage
of the strong beat present in most pop music, and record just one
chroma vector per beat.  This beat-synchronous chromagram
representation represents a typical music track
in a few thousand values, yet when resynthesized back into audio via
modulation of octave-invariant ``Shepard tones'', the melody and
chord sequences of the original music typically remain recognizable
\cite{Ellis2007a}.  To the extent, then, that beat-chroma
representations preserve tonal structure, they
are an interesting domain in which to search for patterns -- rich enough
to generate musically-relevant results, but simplified enough to
abstract away some other aspects of the original audio, such as
instrumentation and other stylistic details.

Specifically, this paper identifies common patterns in beat-synchronous
chromagrams by learning codebooks from a large set of examples.
The individual codewords consist of short beat-chroma patches, typically
between 1 and 8 beats, which are able to represent the entire dataset
of millions of patterns with the minimum error given a small codebook of
a few hundred entries.  Our goal is to identify meaningful information
about the musical structure represented in the entire database by
examining individual entries in this codebook.  Since the common
patterns represent a higher-level description of the musical content,
we also expect them to be useful in other applications, such as
music classification and retrieving tonally-related items.

One major issue in working with small patches concerns temporal
alignment. Learning the same motifs with different offset
(of a few seconds, or audio frames) is redundant. We take advantage
of rhythmic knowledge about the song to align patches with beats.
We can further group beats into bars to produces patches
for meaningful music segments.
The use of the Echo Nest API \cite{EchoNest} for that purpose is 
explained in Subsection \ref{ssec:beatpatch}. Patches built around
those temporal landmarks should capture music motifs more easily.

Prior work using small patches of chroma features includes
the ``shingles'' of \cite{Casey2007}, which they used to
identify ``remixes'' i.e. music based on some of the same
underlying instrument tracks, as well as for matching
performances of Mazurkas \cite{Casey2008}.  That work,
however, was not concerned with extracting the deeper
common patterns underlying different pieces (and did not
use either beat- or bar-synchronous features).  Earlier work in
beat-synchronous analysis includes \cite{Bartsch2001},
who looked for repeated patterns within single songs to
identify the chorus, and \cite{Ellis2007a}, who
cross-correlated beat-chroma matrices to match cover
versions of pop music tracks.  None of these works
examined or interpreted the content of the chroma
matrices in any detail.  In contrast, here we hope to
develop a codebook whose entries are of interest
in their own right.

\begin{figure}[thb]
\begin{center}
\includegraphics[width=.6\columnwidth]{code}
\end{center}
\caption{\small{A typical codeword from a codebook of size $200$ (code
    7 in Figure \ref{fig:codes1}), corresponding to a
    tonic-subdominant chord progression.  The patch is composed of $2$
    bars and the pattern length was set to $8$ beats.  }}
\label{fig:code}
\end{figure}

\section{Approach}\label{sec:approach}
% In this section we review the creation of the beat-chroma patterns using
% the Echo Nest data.  Then we explain the online vector quantization algorithm
% used to learn a codebook of those patterns.

\subsection{Features}
The feature analysis used throughout this work is based on Echo Nest
analyze API \cite{EchoNest}.  
%
For any song uploaded to their platform this analysis returns a chroma
vector (length $12$) for every music event (called ``segment''), and a
segmentation of the song into beats and bars. Beats span multiple
segments, and bars span multiple beats.
%
Averaging the per-segment chroma over beat times results in a
beat-synchronous chroma feature representation similar to that used in
\cite{Ellis2007a}.

% \begin{figure}[htb]
% \begin{center}
% \includegraphics[width=.9\columnwidth]{bar_size_freq}
% \end{center}
% \caption{\small{
% Frequency of the different bar size in the training data.
% }}
% \label{fig:barsize}
% \end{figure}

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
song segmentation than the Echo Nest). 
Practice showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Subsection \ref{ssec:segment}).

%\subsection{Harmonic Patterns}
\subsection{Beat-Chroma Patches} \label{ssec:beatpatch}
% Patterns are defined as a number of beats or bars. In the case of beats,
% we simply cut the set of chromas per beat every $N$ beats. In the case
% of bars, we gather all beats corresponding to the $N$ bars requested.
% The number of chroma vectors we get can vary. We resize the features in the
% time axis to a fixed length. Resizing over clipping can be argued over, 
% but it should
% preserve the main bar characteristics (such as a strong beat at the beginning
% or a note change in the middle).

We use the bar segmentation obtained from the Echo Nest analysis to
break a song into a collection of beat-chroma ``patches'', typically
one or two bars in length.
%
Because the bar length varies, depending on the meter of a particular
song, we resample each patch to a fixed length of typically 4 beats
per bar.  Since the overwhelming majority (82\%) of our training data
consisted of bars that were 4 beats long, this resampling was not
often required.  Most of the remaining bars (10\%) were 3 beats in
length.
% although the majority of the data contains bars that are 4 beats long
The resulting patches consist of $12 \times 4$ or $12 \times 8$ matrices.

Finally, we normalize the patches with respect to the key by rotating
the pattern matrix so that the first row contains the most
energy. This can easily be seen in Figure \ref{fig:code}.
% We considered the same technique for rolling along the other axis and
% be invariant regarding the downbeat. However, it seems less robust to
% noise.
Each patch within a song is normalized independently, so exact
reconstruction of the original song requires knowledge of the
rotation index (key) for each patch.

The representation resulting from this process is invariant to both
the key and meter of the original song.  This enables the study of
broad harmonic patterns found throughout the data, without regard for
the specific musical context.
%
% In the context of clustering, this enables us to study harmonic
% pattern usage without maintaining clusters for the specific context a
% particular patterns occurs in,
In the context of clustering this avoids \eg obtaining separate
clusters for every major triad in duple or triple meter.
% Also, this implies that the pitch envelope is important rather than
% the actual pitches.
%FIXME: Something about why we're not interested in specific chords?
Specific chords do not add much information about a music collection.
A cover song can be performed in a different key, but the motifs it
contains remain the same.

\subsection{Clustering}

% We can then stack beat vectors for a number of bars (usually 1 or 2).

We use an online version of the vector quantization algorithm
\cite{Gersho1991} to cluster the beat-chroma patches described in the
previous section.
% This can also be seen as an online K-means.
For each sample from the data, the algorithm finds the closest cluster
in the codebook and updates the cluster centroid (codeword) to be
closer to the sample based on some learning rate $\ell$.
The clusters are updated as each data point is seen, as opposed to
once per iteration in the standard k-means algorithm.
The details are explained in Algorithm \ref{algo:vq}.
%
As in standard k-means clustering, the codebook is initialized by
choosing $K$ random points from our dataset.
%
Note that this algorithm, although not optimal, scales linearly with
the number of patches seen and can be interrupted at any time
while still resulting in an updated codebook. %not that this is important...
% The setting of the learning rate might become trickier, but this is
% not fundamentally a problem.

\begin{algorithm}
%\caption{Pseudocode of Vector Quantization}
\begin{algorithmic}
\STATE$\ell$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < \ell \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} \dist(p,c)$
\STATE$c \leftarrow c + (p - c) * \ell$
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\caption{\small{Pseudocode for the online vector quantization
    algorithm. Note that we can replace the number of iterations by a
    threshold on the distortion over some test set.}
\label{algo:vq}}
\end{algorithmic}
\end{algorithm}

\section{Experiments}\label{sec:experiments}
In this section we present different clustering experiments. We introduce
our training data and our main testing data. Some detailed settings
of our algorithm are also provided. As for any clustering algorithm, we
measure the influence of the number of codewords and the training set size.

\subsection{Data}
%\subsection{Training data: Cowbell Dataset}
\label{sec:traindata}
%\subsection{Test data: USPOP}
\label{sec:testdata}
Our training data consists of $43,300$ tracks that were uploaded to
morecowbell.dj,\footnote{http://www.morecowbell.dj/} an online service
based on the Echo Nest analyze API which remixes uploaded music by
adding cowbell and other sound effects synchronized in time with the
music.  The $43.3$K songs contain $\mathbf{3.7}$ million non silent
bars which we cluster using the approach described in the previous
section.

For testing, we made use of low quality (32kbps) version of the songs
from the uspop2002 data set \cite{uspop2002}.  This well-known data
contains a wide variety of pop songs, but is not particular otherwise.
% \footnote{We also tried to use Tzanetakis dataset,
% %\cite{Tzanetakis2002a},
% but the $30$ seconds segments seemed hard to analyze by the Echo Nest API, 
% most did not contain any bar.}.
uspop2002 serves as test set to measure how well a codebook learned on
the Cowbell data set can represent new songs.  We obtained Echo Nest features
for the $8651$ songs contained in the dataset.

\subsection{Setting}\label{ssec:setting}
We take one or two bars and normalize the patches to size 4 or 8
respectively.  We learn a codebook of size $K$ over the Cowbell
dataset using the VQ algorithm (Algorithm \ref{algo:vq}). We used a
learning rate of $\ell=0.01$ for $200$ iterations over the whole
dataset.
%
We use the codebook to encode the test set.
Each pattern is encoded with only one code. We can measure the average
distance between a pattern and its encoding. We can also measure the use
of the different codes, \ie the size of the cluster around a code.

We use the average squared Euclidean distance as the distortion
measure between chroma patches.  Given a pattern $p_1$
composed of elements $p_{1ij}$, and a similar pattern $p_2$, the
distance between them is:
\begin{eqnarray}
  \dist(p_1,p_2) = \sum_{i,j} \frac{(p_{1ij} - p_{2ij})^2}{\size(p_1)}
  \label{eq:dist}
\end{eqnarray}
We assume $p_1$ and $p_2$ have the same size.  This is enforced by the
meter normalization procedure described in Section \ref{sec:approach}.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{data_sizes}
\end{center}
\caption{\small{Distortion for a codebook of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a codeword or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $0$ (just initialization) to $500$ thousand. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.
}}
\label{fig:data_sizes}
\end{figure}

\subsection{Basic Properties}
This section present some fundamental results of the clustering algorithm.
None of the result appear surprising,  but they are important for 
reproducing the experiments.
\begin{itemize}
\item Encoding performance improves with increasing training data (Figure
\ref{fig:data_sizes}). Regarding the convergence of the performance,
we can imagine using $500$ or $1$K samples per codeword to learn if the
training set size is problematic.
\item Encoding performance improves with increasing codebook size
(Table \ref{tab:cbsize}). The increase seems significant, and
increasing the size of the codebook in an efficient way
is definitely part of our future work.
\item Larger patterns are more difficult to encode, thus requiring
larger codebooks. See Figure \ref{fig:perbeat}. The increase is steeper
until $4$ beats which usually corresponds to one bar.
\end{itemize}

\begin{table}
\begin{center}
\begin{tabular}{llc}
\toprule
Codebook size & Distortion \\
\midrule
1 & $0.066081$ \\
10 & $0.045579$ \\
50 & $0.038302$ \\
100 & $0.035904$ \\
500 & $0.030841$ \\
\bottomrule
\end{tabular}
\end{center}
\caption{\small{Varying codebook size using $50,000$ training samples.
    We train on patterns corresponding to $1$ bar resized to $4$ beats.
  }}
\label{tab:cbsize}
\end{table}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{encoding_per_beat}
\end{center}
\caption{\small{Encoding larger and larger patterns with a fixed size
codebook. The size of the pattern is defined by the number of beats,
the bar information was not taken into account for this experiment.
}}
\label{fig:perbeat}
\end{figure}

\section{Visualization} \label{sec:visu}
% The goal here is to give an intuition to the reader of what kind of
% typical patterns do we get. We also want to have a feel for what the
% clusters look like, how many boring sustained notes are encoded in the
% codebook, can the codebook describes the complexity of the music space, etc.

\subsection{Codebook}\label{sec:codebook}

We train a codebook of size $200$ over patterns of size $12$x$8$ covering
$2$ bars at a time.  The results shown are on the test set from
Section \ref{ssec:artist}.

The 25 most frequently used codewords in the test set are shown in
Figure \ref{fig:codes1}.  
The frequency of use of these codewords are shown in Figure \ref{fig:freqs}.
The codewords primarily consist of sustained
notes and simple chords.  Since they are designed to be key-invariant,
specific notes do not appear.  Instead the first 7 codewords
correspond to a single note sustained across two bars (codeword 0),
perfect fifth (codewords 1 and 2) and fourth intervals (codewords 3
and 6), and a
major and minor triads (codewords 5 and 4, respectively).  Many of the remaining
codewords correspond to common transitions from one chord to another,
\eg a tonic-subdominant transition in codes 7 and 9
(Cmaj
$\rightarrow$ Fmaj or C5 $\rightarrow$ F5 as a guitar power chord)
% assuming that chroma 0 is C) 
and the reverse tonic-dominant transition in code 21 (\ie Fmaj
$\rightarrow$ Cmaj).
%FIXME: Does this transition have a better name like ``relative x''?
% C5 -> F5 in 'guitar chords', basically a power chord

\begin{figure}[h]
\begin{center}
\includegraphics[width=.9\columnwidth]{codes1}
\end{center}
\caption{\small{
25 most frequently used codes in the codebook %in term of frequency of use
in the artist20 test set. 
This codebook contains $200$ codes trained on $2$ bar, $12 \times 8$
patches.
The numbers 
%above each panel correspond to the display are the index (from most used to least) and
in parentheses correspond to 
the percentage of use in the test set.
%There are $71832$ patterns in total in the dataset.
}}
\label{fig:codes1}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.99\columnwidth]{freqs}
\end{center}
\caption{\small{Usage frequency of the patterns in the codebook from
Figure \ref{fig:codes1} on the artist20 dataset.  There are $71,832$
patterns in total in the dataset. Also frequency on the training set.
Note that even though all codewords are initialized from sample, some
are used only once (training set) or never (test set).
}}
\label{fig:freqs}
\end{figure}

Another interesting visualization uses LLE\footnote{implementation:
  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}
algorithm \cite{Roweis2000} to display codewords arranged in
coherent groups. See Figure \ref{fig:lle}. For instance, we see
``wideband'' patterns grouping to the right.

Finally, more related to our particular data, we try to assess how many
sustained patterns are represented in the codebook. We look at
the average variance over time in Figure \ref{fig:code_var}.
25\% of the codebook correspond
to very simple sustained patterns similar to the top row of Figure
\ref{fig:codes1}. For $2$ bars codewords, it seems that patterns
containing more complex temporal activity account for only $33$\% of
the data.

Showing codewords of $2$ bars is practical and consistent with our experiments
in Section \ref{sec:exps2}. However, longer codewords can also reveal
interesting structure. A random sample of such codewords is presented
in Figure \ref{fig:codes8bars}. Most used codewords would be less interesting,
as they all are sustained patterns.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{codes_lle}
\end{center}
\caption{\small{
Locally linear embedding (LLE) visualization of the codebook.
LLE tries to keep near neighbors (here according to euclidean distance)
in the same $2$D neighborhood. Shown patterns were randomly selected
in each neighborhood.
}}
\label{fig:lle}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{code_variance}
\end{center}
\caption{\small{Average variance in time of a codebook. The vertical
axis cuts at the $63$rd pattern, a rough estimate of the number of 
codewords consisting of sustained chords.
}}
\label{fig:code_var}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{close_patterns1}
\end{center}
\caption{\small{Cluster around centroid presented in
Figure \ref{fig:code}. Taken from the artist20 dataset, the cluster
size is actually $639$. Shown samples were randomly selected.
}}
\label{fig:cluster}
\end{figure}

\subsection{Cluster}
Visualizing a codebook is half of the story, we also want to understand
clusters, \ie the patterns from actual song that
surround a given centroid (codeword). Unsurprisingly, real patterns are
noisier than codes. However, centroids seem to capture important aspects
of the patterns. We could compare the centroids with the result of
a lowpass filtering. In Figure \ref{fig:cluster} we show elements of the same
cluster, and in Figure \ref{fig:cluster_diff}, we emphasize the difference
with the centroid. We hope for the difference to look uniformly random,
the result seems to fit that assumption.

%FIXME: we want the differences to be close to uniformly random, right
%(hence centroid)?
%And that's kind of what they look like.  Does Fig
%\ref{fig:cluster_diff} really add much?

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{close_patterns_diff}
\end{center}
\caption{\small{First three patterns of figure \ref{fig:cluster}
($2$nd line) presented with the centroid from Figure \ref{fig:code}
($1$st line) and the absolute difference between both ($3$rd line).
}}
\label{fig:cluster_diff}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{codes_8bars}
\end{center}
\caption{\small{Sample of longer codewords spanning over $8$
    bars. Codewords selected randomly from the codewords used in the
    test set. Percentage of use display in parentheses.  Most patterns
    consist of sustained note or chords, but code 0 shows a repeating
    two chord pattern and code 4 contains two bars of a
    1$\rightarrow$1$\rightarrow$1$\rightarrow$2 progression.}}
\label{fig:codes8bars}
\end{figure}

\subsection{Example song encoding}

Figure \ref{fig:encodesong} shows a typical example of encoding a song
using the codebook described in Section \ref{sec:codebook}.  Both the
Echo Nest chroma features for an example song and the corresponding
quantization are displayed.  The quantized representation retains
dominant harmonic structure of the original features, but is smoothed
in time due to the use of 2 bar long codewords.
%
This is consistent with the codebook described in Section
\ref{sec:codebook} which primarily consists of sustained chords across
4 or 8 beats.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{song_encoded}
\end{center}
\caption{\small{Good Day Sunshine by \textit{The Beatles}.
Original song and its encoding using a $200$ entry codebook learning
$2$ bars at a time.
}}
\label{fig:encodesong}
\end{figure}

\section{Applications}\label{sec:exps2}
In this section we describe two other applications of the clustering
described in this paper: song segmentation, and artist
recognition.  This demonstrates the utility of unsupervised clustering
to supervised tasks. The representation created from large collections
captures meaningful information without being forced to do so.
% Although our algorithm was not devised for any of these
% tasks, but some information contained in the centroids can be
% naturally used.

\subsection{Song Segmentation} \label{ssec:segment}

Since the clustering described in Section \ref{sec:approach} is based
on the segmentation of the signal in to bars, it stands to reason that
the cluster centroids should contain information related to bar
alignment, such as the presence of a strong beat on the first beat.
%
In this section we investigate whether it is possible to recover the
bar segmentation of new songs using the information contained in the
cluster centroids.  The results are shown Table \ref{tab:offset}. We find
the right segment with an accuracy of $62\%$ where random is
$25\%$. Details of the experiment follow.

We train a codebook of size $100$ on bars resized to $4$ beats. Then,
we take the longest sequence of bars of $4$ beats for each song
in the test set.  As
we mentioned before, $4$ beats is the most common bar size using the
Echo Nest segmentation. We then encode each of these sequences using an offset
of $0$, $1$, $2$ or $3$ beats. The best offset for a test sequence
(the one with the lowest distortion)
should correspond to the bar alignment in the original
feature matrix.  The correct answer is an offset of $0$.

%FIXME: Actually I'm (ronw) not sure how this was done.  Every 4 beat
%segment was treated separately?  Its not looking for a per-song alignment?
%TBM: yes, I simply summed the divergence from all patterns for a given
% divergence, the lowest sum is the winner. So it's a per-song alignment.

The result presented here suggest a dynamic programming setting where
we could find the optimal bar segmentation according to our codebook. 
We could try
all possible bar length (in terms of beats) and not restrict ourselves
to $4$ beats bars and offsets between $0$ and $3$. This is part of
our future work.

\begin{table}
\begin{center}
\begin{tabular}{cc}
\toprule
Offset & \% of times chosen \\
\midrule
0 & $\mathbf{62.6}$\\
1 & $16.5$\\
2 & $09.4$\\
3 & $11.5$\\
\bottomrule
\end{tabular}
\end{center}
\caption{\small{
Song segmentation experiment:
best offset according to our codebook. The real one is $0$.
Random guess gives an accuracy of $25$\%.
}}
\label{tab:offset}
\end{table}

\subsection{Artist Recognition} \label{ssec:artist}

We apply our codebook to the artist recognition task.
Our data set encompass $1402$ songs from $20$ artists, 
mostly rock and pop of different subgenres. Results exist in the
literature for that set \cite{Ellis2007}. The best accuracy is
$59\%$, whereas using only chroma as a representation yields an
accuracy of $33\%$. The latter results is obtained through a
Gaussian mixture model.

Once again, the clustering method proposed in this work is not
intended to perform well at artist recognition. An obvious fact is that
we lose all timbral information which would be useful. However, we can
imagine that some artist use more often some patterns than others.
Our accuracy result is $23.4\%$ while random
is around $5\%$. 
The confusion matrix can be seen in Figure \ref{fig:conf_mat}.
Details of the experiment follow.

All songs in the artist dataset are encoded as an histogram of the codes
used in that song. The frequency values are normalized by the number
of patterns in the song. We test each song in a leave-one-out setting.
Leaving aside the test song, we represent each of the $20$ artists by the 
average of
the song histograms of that artist. The test song is compared to
the histogram of each artist using euclidean distance.

\begin{figure}[htb]
%\begin{center}
\includegraphics[width=.9\columnwidth]{conf_mat_per_artist}
%\end{center}
\caption{\small{Confusion matrix for artist recognition task.}}
\label{fig:conf_mat}
\end{figure}

Looking at the most typical patterns per artist can be insightful.
To find these patterns, we compare its use by one artist and divide
by its use by all artists. In Figure \ref{fig:typicalpat} we see 
Metallica's pattern, then Tori Amos and
Suzanne Vega's pattern. These three artists were easily identified.
Artists like Metallica are characterized by ``wideband'' patterns -
featured muddled due to noisy acoustics - lots of transients,
distortion, etc, which can be seen in Figure \ref{fig:metallica2}.
%FIXME: ron, what di you mean: (not by cleaned up by EN's feature analysis)

% ignoring the two following figures
\iffalse
\begin{figure}[htb]
\begin{center}
\includegraphics[width=.4\columnwidth]{metallica_pattern}
\end{center}
\caption{\small{Most typical pattern for Metallica.
COULD BE MIXED WITH TORI AMOS FIGURE.
}}
\label{fig:metallica}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.4\columnwidth]{toriamos_suzannevega_pattern}
\end{center}
\caption{\small{Most typical pattern for Tori Amos and Suzanne Vega.
    Can we show the top 2 or 4 for each to get a better idea?}}
\label{fig:amosvega}
\end{figure}
\fi

\begin{figure}[htb]
  \centering
  \subfloat[][\small{Metallica}]{\label{fig:metallica2}\includegraphics[width=0.32\columnwidth]{metallica_pattern}}
  \hspace{5mm}                
  \subfloat[][\small{Tori Amos / Suzanne Vega}]{\label{fig:amosvega2}\includegraphics[width=0.32\columnwidth]{toriamos_suzannevega_pattern}}
  \caption{\small{Typical patterns for different artists.}}
  \label{fig:typicalpat}
\end{figure}

\section{Conclusion and Future Work}
We presented a practical method to perform large-scale clustering of
harmonic patterns. We assessed the basic properties of the method through
experiments on a large collection of music. We suggested many ways
to interpret the data (such as LLE) and assert the merit of the codebook
through further experiments.
We discussed the possibility to move to even larger scales
and we provided our source code.\footnote{Code not yet released to preserve
submission's anonymity}

As for enhancements, we specifically did not compress our features using
Gaussian mixtures or other generative model. That being said, these methods
could be use to develop better distance measures between patterns.
The use of the euclidean distance is arbitrary. Summarizing patches
with Gaussians, and then comparing the distance between those Gaussians,
could reduce the influence of the noise in the distance measure.

Moving on to larger scales, we could imagine merging codes together and
splitting some on two if the clusters become too small or too large.We could
also cluster codebooks themselves, in a similar fashion as hierarchical
Gaussian mixtures \cite{Vasconcelos2001}.
%  (similar idea used for indexing \cite{Muja2009})

\small
%\section{Acknowledgements}
%Thierry is NSERC graduate fellow, or some title like that.
%Ron has a grant too.
% Graham for the NMF stuff and numerous discussions.

%\begin{thebibliography}{citations}
%\bibitem{Someone:04} 
%X. Someone and Y. Someone:
%{\it Title of the Book},
%Editorial Acme, Utrecht, 2004.
%\end{thebibliography}

\bibliography{tbm_bib}

\end{document}