### 2010-07-22 Paper/Poster/poster.tex
getting there, some feedback from Ron included

%\documentclass[landscape,a0b,final]{a0poster}
\documentclass[portrait,a0b,final]{a0poster}

\usepackage{epsfig}
\usepackage{multicol}
\usepackage{pstricks,pst-grad}
\usepackage{url}
\usepackage{wrapfig}
\usepackage{algorithm,algorithmic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Definition of some variables and colors
\setlength{\columnsep}{3cm}
\setlength{\columnseprule}{2mm}
\setlength{\parindent}{0.0cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%               Background                     %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\background}[3]{
  \newrgbcolor{cgradbegin}{#1}
  \newrgbcolor{cgradend}{#2}
  \psframe[fillstyle=gradient,gradend=cgradend,
  gradbegin=cgradbegin,gradmidpoint=#3](0.,0.)(1.\textwidth,-1.\textheight)
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                Poster                        %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newenvironment{poster}{
  \begin{center}
  \begin{minipage}[c]{0.98\textwidth}
}{
  \end{minipage} 
  \end{center}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                pcolumn                       %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newenvironment{pcolumn}[1]{
  \begin{minipage}{#1\textwidth}
  \begin{center}
}{
  \end{center}
  \end{minipage}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                pbox                          %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newrgbcolor{lcolor}{00. 0. 0.80}
\newrgbcolor{gcolor1}{1. 1. 1.}
\newrgbcolor{gcolor2}{.80 .80 1.}

\newcommand{\pbox}[4]{
\psshadowbox[#3]{
\begin{minipage}[t][#2][t]{#1}
#4
\end{minipage}
}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                myfig                         %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \myfig - replacement for \figure
% necessary, since in multicol-environment 
% \figure won't work
\newcommand{\myfig}[3][0]{
\begin{center}
  \vspace{1.5cm}
  \includegraphics[width=#3\hsize,angle=#1]{#2}
  \nobreak\medskip
\end{center}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                mytable                       %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \mytable - replacement for \table
% necessary, since in multicol-environment 
% \table won't work
\newcommand{\mytable}[3][0]{
\begin{center}
  \vspace{1.5cm}
  \includegraphics[width=#3\hsize,angle=#1]{#2}
  \nobreak\medskip
\end{center}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                mycaption                     %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \mycaption - replacement for \caption
% necessary, since in multicol-environment \figure and
% therefore \caption won't work

%\newcounter{figure}
\setcounter{figure}{1}
\newcommand{\mycaption}[1]{
  \vspace{0.5cm}
  \begin{quote}
    {{\sc Figure} \arabic{figure}: #1}
  \end{quote}
  \vspace{1cm}
  \stepcounter{figure}
}

%\newcounter{table}
\setcounter{table}{1}
\newcommand{\mytablecaption}[1]{
  \vspace{1.0cm}
  \begin{quote}
    {{\sc Table} \arabic{table}: #1}
  \end{quote}
  \vspace{1cm}
  \stepcounter{table}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\background{1.0 1.0 1.0}{1.0 1.0 1.0}{0.3}

\vspace*{2cm}

\newrgbcolor{lightred}{0.7 0.0 0.0}
\newrgbcolor{white}{1.0 1.0 1.0}
\newrgbcolor{whitepink}{1.0 0.8 0.9}

\begin{poster}

%%%%%%%%%%%%%%%%%%%%%
%%% Header
%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\begin{pcolumn}{0.98}

\makebox[0.957\textwidth]{

%%% Unisiegel
\begin{minipage}[c][7cm][c]{0.2\textwidth}
  \begin{center}
    \includegraphics[width=7cm,angle=0]{labrosa-logo.eps} \\
    \vspace{.5cm}
    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
  \end{center}
\end{minipage}
%\begin{minipage}[c][9cm][c]{0.1\textwidth}
%  \begin{center}
%    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
%  \end{center}
%\end{minipage}

\pbox{0.57\textwidth}{}{linewidth=2mm,framearc=0.3,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
%%% Title
\begin{center}
\begin{minipage}[c][10cm][c]{0.90\textwidth}
  \begin{center}
    {\sc \Huge Clustering Beat-Chroma Patterns}\\[5mm]
  {\sc \Huge in a Large Music Database}\\[10mm]
    {\Large Thierry Bertin-Mahieux $^\dagger$, Ron J. Weiss $^\ddagger$ and Daniel P. W. Ellis $^\dagger$\\[3.5mm]
     \large $\dagger$ LabROSA, Dept. of Electrical Engineering, Columbia University \\
     \large $\ddagger$ Music and Audio Research Laboratory, New York University \\
    \tt{\{thierry,ronw,dpwe\}@ee.columbia.edu}}
  \end{center}
\end{minipage}
\end{center}
}

%%% GK-Logo
\begin{minipage}[c][7cm][c]{0.2\textwidth}
  \begin{center}
    \includegraphics[width=10cm,angle=0]{MARL_logo_high_res.eps} \\
    \vspace{1.8cm}
    %\includegraphics[width=8cm,angle=0]{nyu_music_tech_logo_2_2_bw.eps}
    \includegraphics[width=14cm,angle=0]{NYU_logo.eps}
  \end{center}
\end{minipage}

}
\end{pcolumn}
\end{center}

\vspace*{0.5cm}

%%% Begin of Multicols-Enviroment
\begin{multicols}{3}

%%% Section 1
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Introduction
    \end{center}}
\end{center}

\vspace{1.0cm}

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Can we infer anything about the underlying structure and common
features of commercial pop music by analyzing very large collections
of music audio?
\item Our interest: tonal content of the music -- i.e. the
harmony and melody.
\item Beat-synchronous chromagrams: rich enough
to generate musically-relevant results, simplified enough to
abstract away instrumentation and other stylistic details.
\end{list}

\vspace{-.3cm}
\makebox[\columnwidth]{ % replace framebox by makebox
\hspace{-4cm}
\begin{minipage}[l]{.78\columnwidth}
\myfig[0]{code_axis.ps}{.40}
\end{minipage}
\hspace{-3.5cm}
\begin{minipage}[l]{.4\columnwidth}
Typical %learned 
codeword spanning $2$ bars, each bar
considered to be $4$ beats. Codebook size was $200$.\\

A codewords can also be seen as the centroid of a cluster.
\end{minipage}
} % end framebox / makebox
\vspace{.5cm}

%
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item \textbf{Goal}: identify meaningful information
about the musical structure represented in the entire database by
examining individual entries in this codebook.
\item \textbf{Method}: identify common patterns in beat-synchronous
chromagrams by learning codebooks from a large set of examples.
Individual codewords consist of short beat-chroma patches of
between 1 and 8 beats, optionally aligned to bar boundaries.
\item Prior work: ``shingles'' of \cite{Casey2007}, 
beat-synchronous analysis to identify the chorus by \cite{Bartsch2001},
and cover recognition by \cite{Ellis2007a}.
\end{list}

%%% Section 2
\vspace{1.5cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Audio Features - Echo Nest
    \end{center}}
\end{center}

\vspace{1.0cm}

\textbf{Chromagrams from the Echo Nest online API.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Feature analysis based on Echo Nest analyze API \cite{EchoNest}.
\item For any song, EchoNest provides a chroma
vector (length $12$) for every music event (called ``segment''), and a
segmentation of the song into beats and bars.
\item Averaging per-segment chroma over beat times results in a
beat-synchronous chroma feature representation.
\item Dataset size: $43,000$ songs.
\end{list}

\vspace{.5cm}
\textbf{Beat-Chroma Patches.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item We use Echo Nest analysis to break a song into a collection of 
beat-chroma ``patches'', typically one or two bars in length.
\item $82\%$ of the bars in our data were 4 beats long.
\item Other cases: we resample patches to a fixed length of 4 beats.
\item We rotate patches so that the first row contains the most energy.
This is analogous to key normalization, but
each patch is normalized independently.
\end{list}

%%% Figure example
%\begin{center}
  % first argument: eps-file
  % second argument: stretching-factor relative to Column-width (<1)
  % optional argument: rotation angle (0-360), default=0
  %\myfig[0]{myfig.eps}{0.8}
  %\mycaption{Some caption}
%\end{center}

\myfig[0]{song_encoded.ps}{.7}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
Exemple of an encoded song (Good Day Sunshine
by \textit{The Beatles}) using a codebook. Codebook size:
$200$, codewords: $2$ bars.
\end{small}
\end{minipage}
\end{center}

\myfig[0]{codes_8bars.ps}{.7}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
Longer codewords spanning $8$ bars, randomly selected from
a $100$-entry codebook. Most patterns consist of sustained notes 
or chords, but code 0 shows one-bar alternations between 
two chords, and code 4 contains two cycles of a
$1\rightarrow1\rightarrow1\rightarrow2$ progression.

\end{small}
\end{minipage}
\end{center}

% force a new column
\newpage

%%% Section 3
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Vector Quantization
    \end{center}}
\end{center}

\vspace{1.0cm}

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Vector Quantization algorithm \cite{Gersho1991} to cluster beat-chroma
patches
\item Can be seen as online $k$-means
\item Initialized with $k$ random patches from the data
\item Although not optimal, scales linearly with
the number of patches seen
\end{list}

\vspace{1cm}

\makebox[\columnwidth]{ % replace framebox by makebox

\begin{minipage}[l]{.4\columnwidth}
%\caption{Pseudocode of Vector Quantization}
\begin{algorithmic}
\begin{algorithm}[H]
\vspace{.7cm}
\STATE$\ell$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\REQUIRE $0 < \ell \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE{$c \leftarrow \min_{c \in C_k} \mbox{dist}(p,c)$}
\STATE{$c \leftarrow c + (p - c) * \ell$}
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\vspace{.7cm}
\end{algorithm}
\end{algorithmic}
\end{minipage}

\hspace{1.5cm}
\begin{minipage}[r]{.4\columnwidth}
Intuitively:
\begin{list}{\labelitemi}{\leftmargin=1em \labelsep=1em}
\item For each new patch, find the closest codeword.
\item Bring that codewords closer to the patch by some learning
rate.
\item Iterate.
\end{list}
\end{minipage}

}

%%% Section 4
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Pattern Analysis
    \end{center}}
\end{center}

\vspace{1.0cm}
Additional graphs on distortion, codeword size, etc, can be seen in the paper.
As expected, a larger codebook improves distortion, longer codewords are more
difficult to learn, and a larger training set improves the results.\\

Below are samples from actual codebooks and centroids.

\myfig[0]{close_patterns1.ps}{.8}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
Cluster around first centroid (in \textit{introduction} section).
Random selection from a cluster of size $639$. 
This gives an intuition of the
variance in one cluster.
\end{small}
\end{minipage}
\end{center}

\myfig[0]{codes1.ps}{.8}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
The 25 codes most commonly used. 
Codes are from a $200$-entry codebook trained
on $2$ bars, $12 × 8$ patches. The proportion of patches accounted
for by each pattern is shown in parentheses.
\end{small}
\end{minipage}
\end{center}

\myfig[0]{close_patterns_diff.ps}{.8}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
First three patterns from figure above ($2$nd row) presented
with the centroid ($1$st row) and the absolute difference between 
both ($3$ row).
\end{small}
\end{minipage}
\end{center}

% force a new column
\newpage

\mbox{} % dumb, but need something before vspace
\vspace{-1cm}
\myfig[0]{codes_lle.ps}{.8}

% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
Possible visualization of the codebook using locally linear
embedding (LLE).
\end{small}
\end{minipage}
\end{center}

%%% Section 5
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Experiments
    \end{center}}
\end{center}

\vspace{1.0cm}

\begin{minipage}[c]{\columnwidth}

We present two applications of the beat-chroma codebooks
to illustrate how the ``natural'' structure identified via 
unsupervised clustering can provide useful 
features for subsequent supervised tasks.
\vspace{1cm}

\begin{wrapfigure}{r}{0.6\textwidth}
  \vspace{-20pt}
  \begin{center}
    \includegraphics[width=0.58\textwidth]{conf_mat_per_artist.ps}
  \end{center}
  \vspace{-20pt}
\end{wrapfigure}

\textbf{Artist recognition task}.
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item {\it artist20} data set: $1402$ songs from $20$ artists.
\item Baseline: GMM on MFCC ($59\%$), GMM on chromas ($33\%$),
random ($5\%$).
\item We get an accuracy of $\mathbf{23.4\%}$.
\end{list}
%We use the {\it artist20} data set: $1402$ songs from $20$ artists, 
%mostly rock and pop of different subgenres.  
%Previously published results using GMMs on MFCC features achieve an 
%accuracy of $59\%$, whereas using only chroma as a representation yields an
%accuracy of $33\%$  \cite{Ellis2007}.
%
%We get an accuracy of $\mathbf{23.4\%}$, random baseline 
%is around $5\%$. 
%The confusion matrix is shown here, note 
%that certain artists are recognized at an accuracy far above the average.

\end{minipage}

\vspace{1.5cm}

\makebox[\columnwidth]{ % replace framebox by makebox for no edges

\begin{minipage}[l]{0.21\columnwidth}
  \begin{center}
      \begin{tabular}{cc}
        Offset & \% of times chosen \\ \hline \hline
        0 & $\mathbf{62.6}$\\
        1 & $16.5$\\
        2 & $9.4$\\
        3 & $11.5$
      \end{tabular}
  \end{center}
\end{minipage}

\hspace{5cm}

\begin{minipage}[r]{.5\columnwidth}
\textbf{Bar alignment task}.
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Clustering based on the signal segmented in bars.
\item Codewords should contain bar information (e.g. a strong
first beat).
\item Our encoding find the right alignment $62\%$ of the time.
\end{list}
%Since the clustering described is based
%on the segmentation of the signal in to bars, 
%the codewords should contain information related to bar
%alignment, such as the presence of a strong beat on the first beat.
%It is the case $62\%$ of the times, see paper for details.
\end{minipage}

} % end framebox / makebox

\vspace{1.5cm}

%%% Section 6
\vspace{0cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Conclusion
    \end{center}}
\end{center}

\vspace{1.0cm}

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Practical method for large-scale clustering of tonal patterns.
\item Ways to inspect the resulting codebook.
\item Links to existing MIR tasks.
\item \textbf{Python code available.}
\end{list}

%%% References
%\begin{small}
\bibliographystyle{IEEEtran}
\bibliography{tbm_bib}
%\end{small}

\iffalse
% authors
\vspace{-2cm}
\myfig[0]{authors.eps}{.8}
% figure caption
\vspace{-2cm}
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{center}
\begin{small}
Looking forward to discuss our work with you!
\end{small}
\end{center}
\end{minipage}
\end{center}
\fi

\end{multicols}

% FOOTNOTE
\vspace{1cm}
\begin{minipage}{\textwidth}
\begin{small}
\begin{flushright}
Acknowledgements: thanks to G. Grindlay for
helpful comments. TBM is supported by a
NSERC scholarship. This material is based upon work
supported by IMLS grant LG-06-08-0073-08 and by NSF
grant IIS-0713334. \\
%Any opinions, findings and conclu-
%sions or recommendations expressed in this material are
%those of the authors and do not necessarily reflect the views
%of the sponsors.\\
Poster presented at ISMIR 2010 (Utrecht) for the paper:
Clustering Beat-Chroma Patterns in a Large Music Database by
T. Bertin-Mahieux, R. J. Weiss and D. P. W. Ellis. Contact:
\url{tb2332@columbia.edu}, code: \url{http://www.columbia.edu/~tb2332/}.
\end{flushright}
\end{small}
\end{minipage}

\end{poster}
\end{document}